---
title: "Introduction to the UrbanTempo package"
author: "Timothée Flutre and Marie Flutre"
date: "`r format(Sys.time(), '%d/%m/%Y %H:%M:%S')`"
output:
  rmarkdown::html_vignette:
    toc: true
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: TRUE
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: TRUE
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!--
setwd("~/src/UrbanTempo/vignettes/")

library(devtools)
build_vignettes()

library(rmarkdown)
render("introduction.Rmd", "html_document")
-->


# Preamble

This vignette is available under the [CC BY-SA 4.0](http://creativecommons.org/licenses/by-sa/4.0/) license.

To run it, the external package [UrbanTempo](https://github.com/marieflutre/UrbanTempo) is required:
```{r load_pkg}
library(UrbanTempo)
```

The R chunk below is used to assess how much time it takes to execute the R code in this document (see the appendix at the end):
```{r time_0}
t0 <- proc.time()
```


# Context

TODO


# Show an example with fake data

## Create data

Let's create a minimal data set of two space types, with a total of one interval and two ponctual events over two hours, and save it into a file:
```{r fake_data_create}
tmpd <- tempdir() # temporary directory
input.df <- data.frame("date"=c("2017-04-23", "2017-04-23", "2017-04-23"),
                       "time"=c("07:24", "07:13", "08:17"),
                       "shops"=c("shop end", "shop start", NA),
                       "street"=c("hawker", NA, "hawker"),
                       stringsAsFactors=FALSE)
tmpf <- paste0(tmpd, "/input_temporalities.tsv") # temporary file
write.table(x=input.df, file=tmpf, quote=FALSE, sep="\t",
            row.names=FALSE, col.names=TRUE)
```

## Load data

Such a file can be loaded with the following function, whose output is here assigned to a variable named `tpr`:
```{r fake_data_load}
tpr <- readTemporalities(file=tmpf, duration.event=60, verbose=1)
```

More details are available in the help of this function, via `?readTemporalities`.


## Plot data

First, we need to specify the appearance of each type of temporalities, such as their colors which can be chosen via a palette:

* http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf

* https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/colorPaletteCheatsheet.pdf

```{r fake_data_app}
app <- list("shops"=data.frame(col="chocolate", row.names="shop"),
            "street"=data.frame(col="blueviolet", row.names="hawker"))
```

Then, the input temporalities contained in the `tpr` variable can be plotted with the following function:
```{r fake_data_plot, fig.width=8}
plotTemporalities(temporalities=tpr, appearances=app,
                  main="Fake example of urban temporalities",
                  data.source="fake example, 2017", verbose=1)
```

More details are available in the help of this function, via `?plotTemporalities`.


## Clean

Remove the temporary file:
```{r fake_data_clean}
if(file.exists(tmpf))
  file.remove(tmpf)
```


# Reproduce the figure from Gibert's PhD thesis (2014)

TODO


# Appendix

Below are some details about this vignette:
```{r info}
t1 <- proc.time()
t1 - t0
print(sessionInfo(), locale=FALSE)
```
